---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---

# Introduction

This study is a part of Google Data Analytics Professional Certificate. The goal is to analyze the Divvy bike-share program following the six steps of the data analysis process: ask, prepare, process, analyze, share, and act. The study will take from january of 2024 to january of 2025 as the time frame to analyze the data.

```{r}
## Load libraries
library(tidyverse)
library(DBI)
library(RMySQL)
library(hrbrthemes)
library(viridis)
library(forcats)
library(ggmap)
```

# Ask

Out goal here is answer the following question: *How do annual members and casual riders use Cyclistic bikes dierently?*

For this to happen, we will produce a report with the following deliverables:

1.  A clear statement of the business task
2.  A description of all data sources used
3.  Documentation of any cleaning or manipulation of data
4.  A summary of your analysis
5.  Supporting visualizations and key ndings
6.  Your top three recommendations based on your analysis

To better translate this step into good questions that we will ask of the data provided, we need to think about what problems we are trying to solve and how the insights that emerge will be able to guide business decisions.

In addition, it is essential to understand the expectations of stakeholders and the activities related to the business, only then will we be able to deliver a clear vision of the activities.

# Prepare

You can check all preparation process running `source("DB/data_wrangle.r")` in the console or opening the file in the IDE.

## Sources

The data used in this study is from the Divvy bike-share program in Chicago. The data is available in CSV format and can be downloaded from the following links: [Datasets](https://divvy-tripdata.s3.amazonaws.com/index.html). The data is divided into one file for each month and includes information about the start and end time of the trip, the start and end stations, the duration of the trip, and the user type. I will use SQL to merge all data, so will be easier to analyze the data. All file will be downloaded and saved in the same folder as the script.

## Loading

I initially loaded all the CSV files into separate tables and merged them into a single one. This way, I kept the CSV data already organized in the database as a backup and could work with the processed data in a new table 't_data', no longer depending on separate files to structure the project.

## Cleaning

I cleaned the data by removing duplicates, missing values, and outliers (like trips with less then 1 minutes or more then 24 hours and also trip with started_at bigger or equal ended_at)

## Transform

I created new columns to help me analyze the data, like hour of the day, day of the week, month and duration of the trip in seconds. Also changed data types to make it easier to work with the data and also optimiza the data storage. Also changed stations id and name to have a standard name for all stations.

# Process

With all data ready for analysis, we can start to process the data to answer the question we asked in the first step. On the next steps we will try to answer the question on the Ask phase: *How do annual members and casual riders use Cyclistic bikes dierently?*

On this phase we will analyze the data to understand the behavior of the users, like the time of the day that the users use the bikes, the duration of the trips, the days of the week that the users use the bikes and the most used stations. Is a exploratory data analysis(EDA) to understand the data and the users.

```{r}
# Connect to mySQL
con <- DBI::dbConnect(RMySQL::MySQL(), 
                      dbname = sub(",$", "", Sys.getenv("DB_NAME")),
                      host = sub(",$", "", Sys.getenv("DB_HOST")),
                      username = sub(",$", "", Sys.getenv("DB_USER")),
                      password = sub(",$", "", Sys.getenv("DB_PSWD")))
# Load data
t_data <- DBI::dbGetQuery(con, "SELECT * FROM t_data")

```

## Data Overview

```{r}
# Fix data types
# t_data$started_at <- as.Date(t_data$started_at)
# t_data$ended_at <- as.Date(t_data$ended_at)
## Load Head
head(t_data)
```

## Exploratory Data Analysis

```{r}
# Data manipulation for ABS(Z-SCORE) >= 3 and remove duration < 60 seconds
data <- t_data %>%
  select(rideable_type, member_casual, trip_duration) %>%
  filter(trip_duration > 0) %>% 
  mutate(trip_duration = round(as.numeric(trip_duration)/60, 0)) %>%
  filter(abs(as.vector(scale(as.numeric(trip_duration))))< 3)
```

```{r}
# Group by start station name and count trips
stations <- t_data %>%
  group_by(start_station_name) %>%
  summarise(n_trips = n(), lng = mean(start_lng), lat = mean(start_lat)) %>%
  arrange(desc(n_trips))  %>%
  filter(start_station_name != "")
```

```{r}
# Grouping for timeframe analysis
data_time <- t_data %>%
  select(rideable_type, member_casual, year, month, day_of_week,hour_of_day, trip_duration) %>%
  mutate(day_of_week = factor(day_of_week, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")), month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
  group_by(year, month, day_of_week,hour_of_day, rideable_type, member_casual) %>%
  summarise(n_trips = n(),trip_duration=sum(trip_duration))%>%
  arrange(year, month, day_of_week,hour_of_day, rideable_type, member_casual)
```

### Histogram of number of trips by trip duration



### Histogram of number of trips by trip duration



### Frequecy of trips by rideable type



### Box plot for trip duration



### Percentile of trips by rideable type



### Daily use Analysis



Plot travel

```{r}

",```

### Heatmap of trips on geom_map of chicago

```{r}
# geo data
geo_data <- t_data %>%
  select(rideable_type, member_casual, trip_duration, start_lat, start_lng) %>%
  filter(trip_duration > 0) %>% 
  mutate(trip_duration = round(as.numeric(trip_duration)/60, 0)) %>%
  filter(abs(as.vector(scale(as.numeric(trip_duration))))< 3)
```

```{r}
chi_map <- read_sf("https://raw.githubusercontent.com/thisisdaryn/data/master/geo/chicago/Comm_Areas.geojson")
```

```{r}
geo_data %>%  ggplot() + 
  geom_sf(data = chi_map, fill) + 
  geom_point(aes(x=start_lng, y=start_lat, color=member_casual),alpha = 0.6, size=0.1)+
  scale_fill_viridis(discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  labs(title = "Total hours by Member Type",
       subtitle = "January 2024 to January 2025",
       x = "Type of user",
       y = "Total hours",
       fill = "Rideable Type") +
  theme_minimal()
```

### Total hours spent in each type of rideable separated by member type.

```{r}
data %>% 
  group_by(member_casual, rideable_type) %>% 
  summarise(total_duration = sum(as.numeric(trip_duration)/3600)) %>%
  ggplot(aes(x = member_casual, y = total_duration, fill = rideable_type)) +
  geom_col(alpha = 0.6) +
  scale_fill_viridis(discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  labs(title = "Total hours by Member Type",
       subtitle = "January 2024 to January 2025",
       x = "Type of user",
       y = "Total hours",
       fill = "Rideable Type") +
  theme_minimal()+
  facet_wrap(vars(rideable_type))
```

### Average minutes spent in each type of rideable separated by member type.

```{r}
data %>% 
  group_by(member_casual, rideable_type) %>% 
  summarise(total_duration = mean(as.numeric(trip_duration)/60)) %>%
  ggplot(aes(x = member_casual, y = total_duration, fill = rideable_type)) +
  geom_col(alpha = 0.6) +
  scale_fill_viridis(discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
  labs(title = "Average Minutes by Member Type",
       subtitle = "January 2024 to January 2025",
       x = "Type of user",
       y = "Average of Minutes",
       fill = "Rideable Type") +
  theme_minimal()+
  facet_wrap(vars(rideable_type))
```
