---
title: "Bike-Share_Study"
author: "Pablo Diego Pereira"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

This study is a part of Google Data Analytics Professional Certificate. The goal is to analyze the Divvy bike-share program following the six steps of the data analysis process: ask, prepare, process, analyze, share, and act. The study will focus on the first quarter of 2019 and 2020 to identify any changes in the program's use over time. 

## Prepare

### Data Sources

The data used in this study is from the Divvy bike-share program in Chicago. The data is available in CSV format and can be downloaded from the following links:

Datasets:
 - [Divvy 2019 Q1](https://docs.google.com/spreadsheets/d/1uCTsHlZLm4L7-ueaSLwDg0ut3BP_V4mKDo2IMpaXrk4/template/preview?resourcekey=0-dQAUjAu2UUCsLEQQt20PDA#gid=1797029090)
 - [Divvy 2020 Q1](https://docs.google.com/spreadsheets/d/179QVLO_yu5BJEKFVZShsKag74ZaUYIF6FevLYzs3hRc/template/preview#gid=640449855) 

Or: [Online Index](https://divvy-tripdata.s3.amazonaws.com/index.html)

This study will use the data from the first quarter of 2019 and 2020


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
